### 2020-06-18 report/sections/general/theory.tex
Report: progress on chapter 8
Report: feedback chapter 8

In this chapter, we present
\newtheory, an extension of the
canonical multiparty session type (MPST)
theory with a message routing mechanism.
We build upon the variant of MPST theory introduced
in \cite{LessIsMore}
(which we have also discussed in \cref{subsection:bgmpst}), 
and extend it
with constructs for communication actions
that are routed through a participant.
We introduce the extended syntax for types in 
\cref{section:syntax}, and give semantics of our
syntax extensions in \cref{section:lts}.

\newtheory provide the theoretical basis
for allowing developers to implement
protocols with browser-to-browser interactions
(such as the \tprotocol{Two Buyer} protocol
introduced in \cref{section:twobuyer})
over a server-centric network topology
which uses WebSocket transport.
We define a router-parameterised encoding
for MPST into \newtheory in \cref{section:encoding}
and prove the preservation of well-formedness
and communication.
We proceed to extend \codegen to implement \newtheory
in \cref{chap:impl}.

@@

@@

>>>>

In this chapter, we present
\newtheory, an extension of the
canonical multiparty session type (MPST)
theory with a message routing mechanism.
We build upon the variant of MPST theory introduced
in \cite{LessIsMore}
(which we have also discussed in \cref{subsection:bgmpst}), 
and extend it
with constructs for communication actions
that are routed through a participant.

\newtheory provide the theoretical basis
for allowing developers to implement
protocols with browser-to-browser interactions
(such as the \tprotocol{Two Buyer} protocol
introduced in \cref{section:twobuyer})
over a server-centric network topology
which uses WebSocket transport.
We introduce the extended syntax for types in 
\cref{section:syntax}, and give semantics of our
syntax extensions in \cref{section:lts}.
We define a router-parameterised encoding
for MPST into \newtheory in \cref{section:encoding}
and prove the preservation of well-formedness
and communication.
We proceed to extend \codegen to implement \newtheory
in \cref{chap:impl}.

@@

\input{sections/general/theory/soundcomplete}

@@

\section{Summary}
We have presented \newtheory, a variant of
the canonical MPST theory to express \textit{routed communication}.
We introduced extensions to syntax and semantics, and proved
that our extended semantics preserve deadlock-freedom
for well-formed protocols. We defined an encoding from the
canonical theory onto \newtheory, and proved the preservation
of well-formedness and communication.

### 2020-06-18 report/sections/general/theory/encoding.tex
Report: minor fixes to chapter 8
Report: feedback chapter 8

We present an encoding from canonical MPST theory
to \newtheory.
This encoding is \textit{parameterised} by the router role 
(conventionally denoted as $\mrole{s}$);
the intuition is that we encode all communication interactions
to involve $\mrole{s}$.
For example, as motivated in \cref{chap:p2p},
we can encode the \tprotocol{Two Buyer} protocol
into routed multiparty session types with respect to the
\trole{S}eller role as the router.
When we extend \codegen to implement \newtheory in \cref{chap:impl},
the developer can implement the \tprotocol{Two Buyer} protocol
as an interactive web application over WebSocket transport.

@@

Readers may observe striking similarities between
the encoding on global types and the definition of a
centroid in routed communication (\cref{def:centroid}).
In fact, the two are the same -- the encoding on global
types simply expresses the centroid predicate as a function.
We arrive at \cref{lem:enccenter} -- the proof is straightforward
by induction on global types and left as an exercise to the reader.

@@

@@

Similarly, if the original communication does not involve
the router from a local perspective, the local endpoint
will carry out the communication through the router
$\trole{s}$ as a proxy in the encoded local type.
This is expressed in both \rulename{Enc-L-Sel}
and \rulename{Enc-L-Bra}.

@@

\begin{lemma}
\[
\forall G_1, G_2, \mrole{r}, \mrole{s}. ~ \left(
(\proj{G_1}{r}) = (\proj{G_2}{r}) \wedge \mrole{r} \neq \mrole{s}
	\Longrightarrow
\proj{\enc{G_1}{s}}{r} = \proj{\enc{G_2}{s}}{r}
\right)
\]

\label{lem:encprojeq}
\end{lemma}

\begin{proof} By consequence from \cref{lem:enclink}.

Take $G_1, G_2, \mrole{r}, \mrole{s}$ arbitrarily. 
Assume $(\proj{G_1}{r}) = (\proj{G_2}{r})$ and $\mrole{r} \neq \mrole{s}$.

We need to show $\proj{\enc{G_1}{s}}{r} = \proj{\enc{G_2}{s}}{r}$,
but by \cref{lem:enclink}, it is sufficient to show
\[
\enclocal{\proj{G_1}{r}}{r}{s} = \enclocal{\proj{G_2}{r}}{r}{s}
\]

Define an ``inner'' function $f(z) = \enclocal{z}{r}{s}$.

By definition, $\forall x, y. (x = y \Longrightarrow f(x) = f(y))$.

We have $(\proj{G_1}{r}) = (\proj{G_2}{r})$ by assumption,
so we can conclude.

\begin{align*}
(\proj{G_1}{r}) = (\proj{G_2}{r}) 
&\Longrightarrow f(\proj{G_1}{r}) = f(\proj{G_2}{r})  \\
&\Longrightarrow
\enclocal{\proj{G_1}{r}}{r}{s} = \enclocal{\proj{G_2}{r}}{r}{s}
\end{align*}

%\[
%(\proj{G_1}{r}) = (\proj{G_2}{r}) 
%\Longrightarrow
%f(\proj{G_1}{r}) = f(\proj{G_2}{r}) 
%\Longrightarrow
%\enclocal{\proj{G_1}{r}}{r}{s} = \enclocal{\proj{G_2}{r}}{r}{s}
%\]

@@

>>>>

We present an encoding from canonical MPST theory
to \newtheory.
This encoding is \textit{parameterised} by the router role 
(conventionally denoted as $\role{s}$);
the intuition is that we encode all communication interactions
to involve $\role{s}$.
For example, as motivated in \cref{chap:p2p},
we can encode the \tprotocol{Two Buyer} protocol
into routed multiparty session types with respect to the
\trole{S}eller role as the router.
When we extend \codegen to implement \newtheory in \cref{chap:impl},
the developer can implement the \tprotocol{Two Buyer} protocol
as an interactive web application over WebSocket transport.

@@

Readers may observe striking similarities between
the encoding on global types and the definition of a
centroid in routed communication (\cref{def:centroid}).
In fact, the two are the same -- the encoding on global
types simply expresses the centroid predicate as a function.
We arrive at \cref{lem:enccenter} -- the proof is straightforward
by induction on global types.

@@

\begin{proof}
By induction on the structure of $G$.
\end{proof}

@@

Similarly, if the original communication does not involve
the router from a local perspective, the local endpoint
will carry out the communication through the router
$\trole{s}$ in the encoded local type.
This is expressed in both \rulename{Enc-L-Sel}
and \rulename{Enc-L-Bra}.

@@

@@

We motivated the risk of over-serialising communication
with a naive routing definition in \cref{section:routerchallenges}.
We revisit the example and show that our encoding
preserves valid traces in the original communication.

\begin{example}[Encoding Preserves Semantics]
Consider the global type
\[
G = \gcommone{p}{q}{\tmsg{M1}} . ~
	\gcommone{s}{q}{\tmsg{M2}} . ~\tend
\]

We apply our encoding with respect to the router role $\role{s}$
\[
\enc{G}{s} = \grouteone{p}{q}{s}{\tmsg{M1}}.~
	\gcommone{s}{q}{\tmsg{M2}}. ~ \tend
\]

Recall that $l = \aout{s}{q}{\tmsg{M2}}$ can reduce $G$.

\begin{prooftree}
\AxiomC{}
\RightLabel{\rulename{Gr1}}
\UnaryInfC{$
\treduce
	{\left(\gcommone{s}{q}{\tmsg{M2}}.~ \tend\right)}
	{\tend}
	{l}
$}
\AxiomC{$\subj{l} = \mrole{s} \notin \{ \mrole{p}, \mrole{q} \}$}
\RightLabel{\rulename{Gr4}}
\BinaryInfC{$
\treducelong
	{\left(\gcommone{p}{q}{\tmsg{M1}} . ~
		\gcommone{s}{q}{\tmsg{M2}} . ~\tend\right)}
	{\left(\gcommone{p}{q}{\tmsg{M1}} . ~
	\gtransone{s}{q}{\tmsg{M2}} . ~\tend\right)}
	{l}
$}
\end{prooftree}

We show that $\enc{l}{s} = l$ can reduce $\enc{G}{s}$.

\begin{prooftree}
\AxiomC{}
\RightLabel{\rulename{Gr8}}
\UnaryInfC{$
\treduce
	{\left(\gcommone{s}{q}{\tmsg{M2}}.~ \tend\right)}
	{\tend}
	{l}
$}
\AxiomC{$\subj{l} = \mrole{s} \notin \{ \mrole{p}, \mrole{q} \}$}
\RightLabel{\rulename{Gr4}}
\BinaryInfC{$
\treducelong
	{\left(\grouteone{p}{q}{s}{\tmsg{M1}} . ~
		\gcommone{s}{q}{\tmsg{M2}} . ~\tend\right)}
	{\left(\grouteone{p}{q}{s}{\tmsg{M1}} . ~
	\gtransone{s}{q}{\tmsg{M2}} . ~\tend\right)}
	{l}
$}
\end{prooftree}

$\aout{s}{q}{\tmsg{M2}}$ is a prefix of a valid
execution trace for $G$, given below.

\[
\treducelong
	{\treducelong
		{\treducelong
			{\treducelong
				{G}
				{}
				{\aout{s}{q}{\tmsg{M2}}}}
			{}
			{\aout{p}{q}{\tmsg{M1}}}}
		{}
		{\ain{p}{q}{\tmsg{M1}}}}
	{\tend}
	{\ain{s}{q}{\tmsg{M2}}}
\]

Interested readers can verify
that the encoded trace (given below) 
is a valid execution trace for
$\enc{G}{s}$.

\[
\treducelong
	{\treducelong
		{\treducelong
			{\treducelong
				{\enc{G}{s}}
				{}
				{\aout{s}{q}{\tmsg{M2}}}}
			{}
			{\via{s}{\aout{p}{q}{\tmsg{M1}}}}}
		{}
		{\via{s}{\ain{p}{q}{\tmsg{M1}}}}}
	{\tend}
	{\ain{s}{q}{\tmsg{M2}}}
\]

\end{example}

### 2020-06-18 report/sections/general/theory/lts.tex
Report: fix notation
Report: feedback chapter 8

We define labelled transition system (LTS) semantics
over global types (\cref{subsection:newltsglobal}) 
and local types (\cref{subsection:newltslocal})
for \newtheory,
building upon the work of Deni\'elou and Yoshida in
\cite{characterisation}.
We show the soundness and completeness of projection
with respect to the LTSs through proving the
\textit{trace equivalence} of
a global type and the collection of local types projected
from the global type (\cref{subsection:newtraceeq}).
We then use this result to
conclude that \newtheory provide the same
communication safety guarantees from canonical
MPST theory for well-formed global types, 
namely deadlock freedom (\cref{subsection:newdeadlockfreedom}).

@@

The LTS semantics presented in \cite{characterisation}
models \textit{asynchronous communication},
which is consistent with our proposal.
In order to define LTS over global types for
asynchronous communication, we need to
represent intermediate states (i.e. messages in transit)
within the grammar of global types.

\cite{characterisation} added the construct
{$\gtrans{p}{q}{j}{l_i: G_i}{i \in I}$}
to represent that the message $l_j$ has been
sent by $\role{p}$ but not yet received by $\role{q}$.

We add a similar construct
{$\gtransroute{p}{q}{s}{j}{l_i: G_i}{i \in I}$}
to represent that the message $l_j$ has
been sent from $\role{p}$ to the router $\role{s}$
but not yet routed to $\role{q}$.

Because routed communication is treated differently
from normal send and receive actions, the notion
of asynchrony differs between the two types of communication
too. This definition allows us to extend
the LTS semantics defined in \cite{characterisation}
more naturally.

We define the LTS semantics 
over global types,
defined by $\treduce{G}{G'}{l}$,
in \cref{fig:newglobal}.
We highlight and explain the new rules.

@@

We define the LTS semantics 
over local types,
defined by $\treduce{T}{T'}{l}$,
in \cref{fig:newlocal}.
We highlight and explain the new rules.

@@

\item \rulename{Lr4} and \rulename{Lr5} are
analogue to \rulename{Lr1} and \rulename{Lr2}
for sending and receiving messages respectively.
The exception is that the new rules pattern-match
on the proxy role $\role{s}$ on the local type
and the routed label.

@@

\begin{align*}
G_1 &= \gcommone{s}{r}{\tmsg{M1}} \grouteone{r}{q}{s}{\tmsg{M2}} \tend \\
\proj{G_1}{s} &= \tselone{r}{\tmsg{M1}}. ~ 
	\routerone{r}{q}{\tmsg{M2}}{\tend} \\~\\
G_2 &= \gcommone{s}{r}{\tmsg{M1}} \grouteone{p}{r}{s}{\tmsg{M2}} \tend \\
\proj{G_2}{s} &= \tselone{r}{\tmsg{M1}}. ~ 
	\routerone{p}{r}{\tmsg{M2}}{\tend} \\
\end{align*}

>>>>

We define the labelled transition system (LTS) semantics
over global types (\cref{subsection:newltsglobal}) 
and local types (\cref{subsection:newltslocal})
for \newtheory,
building upon the work of Deni\'elou and Yoshida in
\cite{characterisation}.
We show the soundness and completeness of projection
with respect to the LTSs through proving the
\textit{trace equivalence} of
a global type and the collection of local types projected
from the global type (\cref{subsection:newtraceeq}).
We then use this result to
conclude that \newtheory provide the same
communication safety guarantees from canonical
MPST theory for well-formed global types, 
namely deadlock freedom (\cref{subsection:newdeadlockfreedom}).

@@

The LTS semantics presented in \cite{characterisation}
models \textit{asynchronous communication},
which is consistent with our proposal.
In order to define LTS over global types for
asynchronous communication, we need to
represent intermediate states (i.e. messages in transit)
within the grammar of global types.
Deni\'elou and Yoshida \cite{characterisation}
added the construct
{$\gtrans{p}{q}{j}{l_i: G_i}{i \in I}$}
to represent that the message $l_j$ has been
sent by $\role{p}$ but not yet received by $\role{q}$.

We add a similar construct
{$\gtransroute{p}{q}{s}{j}{l_i: G_i}{i \in I}$}
to represent that the message $l_j$ has
been sent from $\role{p}$ to the router $\role{s}$
but not yet routed to $\role{q}$.
We extend the projection operator to 
support this new construct.

\[
\proj{\gtransroute{p}{q}{s}{j}{l_i: G_i}{i \in I}}{r} = \begin{cases}
\tbraproxy{p}{s}{l_i: \proj{G_i}{r}}{i \in I} 
	& \text{if } \mrole{r} = \mrole{q} \\
\routertrans{p}{q}{j}{l_i: \proj{G_i}{r}}{i \in I}
	& \text{if } \mrole{r} = \mrole{s} \\
\proj{G_j}{r}
	& \text{otherwise} \\
\end{cases}
\]

Because routed communication is treated differently
from direct send and receive actions, the notion
of asynchrony differs between the two types of communication
too. This definition allows us to extend
the LTS semantics from \cite{characterisation}
more naturally.

We define the LTS semantics 
over global types,
denoted by $\treduce{G}{G'}{l}$,
in \cref{fig:newglobal}.
We highlight and explain the new rules.

@@

The LTS semantics over local types
need to be extended for routed communication.
Because the router holds a ``global perspective''
on the routed communication, we also need to
represent the intermediate state (at which the message is 
being routed by the router) within the grammar
of local types.
Similar to \cref{subsection:newltsglobal}, we add 
the construct 
$\routertrans{p}{q}{j}{l_i: T_i}{i \in I}$
to represent that, from the local perspective of
the router, the message $l_j$
has been received from $\role{p}$ but not yet routed
to $\role{q}$.

We define the LTS semantics 
over local types,
denoted by $\treduce{T}{T'}{l}$,
in \cref{fig:newlocal}.
We highlight and explain the new rules.

@@

\item \rulename{Lr4} and \rulename{Lr5} are
analogue to \rulename{Lr1} and \rulename{Lr2}
for sending and receiving messages respectively.
The exception is that the new rules pattern-match
on the router role $\role{s}$ on the local type
and the routed label.

@@

\begin{align*}
G_1 &= \gcommone{s}{r}{\tmsg{M1}}.~ \grouteone{r}{q}{s}{\tmsg{M2}}.~ \tend \\
\proj{G_1}{s} &= \tselone{r}{\tmsg{M1}}. ~ 
	\routerone{r}{q}{\tmsg{M2}}{\tend} \\~\\
G_2 &= \gcommone{s}{r}{\tmsg{M1}}.~ \grouteone{p}{r}{s}{\tmsg{M2}}.~ \tend \\
\proj{G_2}{s} &= \tselone{r}{\tmsg{M1}}. ~ 
	\routerone{p}{r}{\tmsg{M2}}{\tend} \\
\end{align*}

### 2020-06-18 report/sections/general/theory/syntax.tex
Report: feedback chapter 8

\section{Syntax}
\label{section:syntax}

We extend the syntax for global types
in \cref{subsection:newglobal} and
local types in \cref{subsection:newlocal}.
The same conventions introduced in \cref{subsection:bgmpst}
apply here: 
we omit message payload types,
and deterministic send and receive actions
are represented by single selection and
single branching constructs respectively.
With these syntax extensions, we appropriately extend
the definition of \textit{projection} and 
\textit{well-formedness} in 
\cref{subsection:newprojection,subsection:newwf}
respectively.

@@

As motivated by \cref{section:routerchallenges},
we need to handle \textit{routed communication} separately
from normal send and receive actions.
We define the syntax of global types for \newtheory 
in \cref{fig:newsyntaxglobal}, and explain the 
new construct for expressing routed communication.
Global types range over $G, G', G_i, \dots$

@@

The new construct $\groute{p}{q}{s}{l_i: G_i}{i \in I}$ reads, 
\textit{
``the communication $\gcomm{p}{q}{l_i: G_i}{i \in I}$
is routed by $\role{s}$''
}.
To be more verbose,
$\role{p}$ offers $\role{q}$ a set of choices
$\left\{l_i\right\}_{i \in I}$,
$\role{q}$ sends their choice to
$\role{s}$, so $\role{p}$ receives the
selection $l_i$ made by $\role{q}$ via $\role{s}$,
and the communication system proceeds with continuation $G_i$.
We refer to $\role{s}$ hereafter as the \textit{router}.
$\role{s}$ ranges over the set of roles 
$\role{p},\role{q},\role{r},\role{s},\dots$,
but we use $\role{s}$ by convention as the server endpoint
is usually the router.
Just as it is assumed that $\role{p} \neq \role{q}$ for
direct communication,
we also assume that
$\role{p} \neq \role{q} \neq \role{s}$ for routed communication.

We also extend the definition of the participants function,
$\pt{G}$, in \cref{def:newpt}, to formalise that the router 
participates in the routed communication.

@@

Local types are obtained from global types via
\textit{projection}, so we also need new
constructs to express routed communication
from the \textit{local} perspective of individual
endpoints.
We explain the extensions to
projection in \cref{subsection:newprojection}.

We define the syntax of local types for \newtheory 
in \cref{fig:newsyntaxlocal}, and walk through
the new constructs below from the perspective of
some arbitrary role $\role{r}$.
Local types range over $T, T', T_i, \dots$

\begin{figure}[!h]
\doublespacing
\[
\begin{array}{rlr}
T ::= &  & \text{Local Types} \\
\mid & \tend & \text{Termination} \\
\mid & \trecvar & \text{Type Variable} \\
\mid & \trec{T} & \text{Recursive Type} \\
\mid & \tsel{p}{l_i: T_i}{i \in I} & \text{Selection} \\
\mid & \tbra{p}{l_i: T_i}{i \in I} & \text{Branching} \\
\mid & \router{p}{q}{l_i: T_i}{i \in I}
& \text{\hl{Routing Communication}} \\
\mid & \tselproxy{p}{q}{l_i: T_i}{i \in I}
& \text{\hl{Proxy Selection}} \\
\mid & \tbraproxy{p}{q}{l_i: T_i}{i \in I}
& \text{\hl{Proxy Branching}} \\
\end{array}
\]
\singlespacing
\captionof{figure}{Global Types in \newtheory}
\label{fig:newsyntaxlocal}
\end{figure}

@@

\item \textbf{Proxy selection:}

If $\role{r}$ has local type $\tselproxy{p}{q}{l_i: T_i}{i \in I}$,
$\role{r}$ is making a selection from the
\textit{intended recipient} $\role{p}$,
but the selection is sent to the intermediate proxy $\role{q}$. 
$\role{r}$ sends their internal choice $l_i$ to $\role{q}$
and proceeds with continuation $T_i$.

\item \textbf{Proxy branching:}

If $\role{r}$ has local type $\tbraproxy{p}{q}{l_i: T_i}{i \in I}$,
$\role{r}$ is offering a choice to the
\textit{intended sender} $\role{p}$,
but the message is received from the intermediate proxy $\role{q}$.
$\role{r}$ receives the external choice $l_i$ from $\role{q}$
and proceeds with continuation $T_i$.

@@

As a general comment, the new local types of proxy selection
and proxy branching should ``behave'' as normal selection and
receive types with respect to their intended recipient and sender
respectively. We keep track of the proxy role in the syntax
to allow us to distinguish between routing communications
from normal send and receive interactions.

@@

We extend the projection operator to be defined
on routed communication. Given the explanations of the
new routing constructs for both global and local types,
\cref{def:newprojection} should be self-explanatory.

\begin{definition}[Projection]
The projection of $G$ onto $\role{r}$,
written $\proj{G}{r}$, is defined as:

@@

As shown in the last (4th) case, 
the same concept of \textit{merging}
applies for routed communication:
when projecting a routed communication
onto a non-participant, the projections of
all continuations must be ``compatible'',
namely they can be merged using the
the \textit{merging operator}, $\mergeop$.
As the merging operator is defined on local
types, we extend the merging operator to be defined
on the extended syntax in \cref{def:newmerge}.

\begin{definition}[Merging Operator]
The merging operator $\mergeop$ on local types
is:

@@

\tmerge
{(\tbraproxy{p}{q}{l_i: T_i}{i \in I})}
{(\tbraproxy{p}{q}{l_j: T'_j}{j \in J})}
	& = & \tbraproxy{p}{q}{l_k: T''_k}{k \in I \cup J} \\
\text{where} & & T''_k = \begin{cases}
T_k & \text{if} ~ k \in I \setminus J \\
T'_k & \text{if} ~ k \in J \setminus I \\
\tmerge{T_k}{T'_k} & \text{if} ~ k \in I \cap J \\
\end{cases} \\
\end{array}
\]
\singlespacing

@@

Recall that proxy selection and proxy
branching behave in the same way as their ``non-proxified''
counterparts -- the merging operator reflects this
similarity.

@@

Recall that well-formedness is a predicate defined 
\textit{solely} on the global type
in canonical MPST theory: a global type $G$ is well-formed
if a projection exists for all its participants.

@@

We need to define the characteristics that $\role{s}$
must display in $G$ to prove that it is the \textit{only}
router. 
We formalise this as an \textit{inductively} defined relation,
$\centroid{G}{s}$, which reads \textit{``$\role{s}$ is the
centroid in $G$''}.
The intuition is that $\role{s}$ is at the centre
of all communication interactions.
We define what it means to be a \textit{centroid} in
\cref{def:centroid}.

@@

\begin{itemize}
\item 
\rulename{$\centroidop$-End}, 
\rulename{$\centroidop$-RecVar}, 
\rulename{$\centroidop$-Rec}:
Trivial.

\item 
\rulename{$\centroidop$-Comm}:
For normal communication, $\role{s}$ must
be a participant and the centroid of all continuations.

@@

>>>>

\section{Syntax for Global and Local Types}
\label{section:syntax}

We extend the syntax for global types
in \cref{subsection:newglobal} and
local types in \cref{subsection:newlocal}
to express routed communication.
The same conventions introduced in \cref{subsection:bgmpst}
apply here: 
we omit message payload types,
and deterministic send and receive actions
are represented by single selection and
single branching constructs respectively.
With these syntax extensions, we appropriately extend
the definition of \textit{projection} and 
\textit{well-formedness} in 
\cref{subsection:newprojection,subsection:newwf}
respectively.

@@

Global types range over $G, G', G_i, \dots$
As motivated by \cref{section:routerchallenges},
we need to handle \textit{routed communication} separately
from normal send and receive actions.
We define the syntax of global types for \newtheory 
in \cref{fig:newsyntaxglobal}, and explain the 
new construct for expressing routed communication.

@@

The new construct $\groute{p}{q}{s}{l_i: G_i}{i \in I}$ reads, 
\textit{
``the communication interaction 
from $\role{p}$ to $\role{q}$
is routed by $\role{s}$''
}.
To be more verbose,
$\role{p}$ offers $\role{q}$ a set of choices
$\left\{l_i\right\}_{i \in I}$,
$\role{q}$ sends their choice to
$\role{s}$, so $\role{p}$ receives the
selection $l_i$ made by $\role{q}$ via $\role{s}$,
and the communication system proceeds with continuation $G_i$.
We refer to $\role{s}$ hereafter as the \textit{router}.
$\role{s}$ ranges over the set of roles 
$\role{p},\role{q},\role{r},\dots$,
but we use $\role{s}$ by convention as the server endpoint
is usually the router.
Just as it is assumed that $\role{p} \neq \role{q}$ for
direct communication,
we also assume that
$\role{p} \neq \role{q} \neq \role{s}$ for routed communication.

We inherit the definitions of the participants function, $\pt{G}$,
introduced in \cref{table:pt}; we extend it in \cref{def:newpt}
to formalise that the router 
participates in the routed communication.

@@

Local types range over $T, T', T_i, \dots$
Local types are obtained from global types via
\textit{projection}, so we also need new
constructs to express routed communication
from the \textit{local} perspective of individual
endpoints.
We explain the extensions to
projection in \cref{subsection:newprojection}.

We define the syntax of local types for \newtheory 
in \cref{fig:newsyntaxlocal}, and walk through
the new constructs below from the perspective of
some arbitrary role $\role{r}$.

\begin{figure}[!h]
\doublespacing
\[
\begin{array}{rlr}
T ::= &  & \text{Local Types} \\
\mid & \tend & \text{Termination} \\
\mid & \trecvar & \text{Type Variable} \\
\mid & \trec{T} & \text{Recursive Type} \\
\mid & \tsel{p}{l_i: T_i}{i \in I} & \text{Selection} \\
\mid & \tbra{p}{l_i: T_i}{i \in I} & \text{Branching} \\
\mid & \router{p}{q}{l_i: T_i}{i \in I}
& \text{\hl{Routing Communication}} \\
\mid & \tselproxy{p}{q}{l_i: T_i}{i \in I}
& \text{\hl{Routed Selection}} \\
\mid & \tbraproxy{p}{q}{l_i: T_i}{i \in I}
& \text{\hl{Routed Branching}} \\
\end{array}
\]
\singlespacing
\captionof{figure}{Global Types in \newtheory}
\label{fig:newsyntaxlocal}
\end{figure}

@@

\item \textbf{Routed slection:}

If $\role{r}$ has local type $\tselproxy{p}{q}{l_i: T_i}{i \in I}$,
$\role{r}$ is making a selection from the
\textit{intended recipient} $\role{p}$,
but the selection is sent to the intermediate router $\role{q}$.,
instead of to $\role{p}$ directly.
$\role{r}$ sends their internal choice $l_i$ to $\role{q}$
and proceeds with continuation $T_i$.

\item \textbf{Routed branching:}

If $\role{r}$ has local type $\tbraproxy{p}{q}{l_i: T_i}{i \in I}$,
$\role{r}$ is offering a choice to the
\textit{intended sender} $\role{p}$,
but the message is received from the intermediate router $\role{q}$,
instead of from $\role{p}$ directly.
$\role{r}$ receives the external choice $l_i$ from $\role{q}$
and proceeds with continuation $T_i$.

@@

As a general comment, the new local types of routed selection
and routed branching should ``behave'' as normal selection and
receive types with respect to their intended recipient and sender
respectively. We keep track of the router role in the syntax
to allow us to distinguish between routing communications
from normal send and receive interactions.

@@

In \cref{def:newprojection},
we extend the projection operator
to be defined
on routed communication.

\begin{definition}[Projection]
The projection of $G$ onto $\role{r}$,
written $\proj{G}{r}$, is extended as:

@@

As shown in the last (4th) case, 
the same concept of \textit{merging}
applies for routed communication:
when projecting a routed communication
onto a non-participant, the projections of
all continuations must be ``compatible'',
namely they can be merged using the
the \textit{merging operator}, $\mergeop$.
As the merging operator is defined on local
types, we extend the merging operator 
(introduced in \cref{fig:merge}) to be defined
on the extended syntax in \cref{def:newmerge}.

\begin{definition}[Merging Operator]
The merging operator $\mergeop$ on local types
is extended as:

@@

\tmerge
{(\tbraproxy{p}{q}{l_i: T_i}{i \in I})}
{(\tbraproxy{p}{q}{l_j: T'_j}{j \in J})}
	& = & \tbraproxy{p}{q}{l_k: T''_k}{k \in I \cup J} \\
\text{where} & & T''_k = \begin{cases}
T_k & \text{if} ~ k \in I \setminus J \\
T'_k & \text{if} ~ k \in J \setminus I \\
\tmerge{T_k}{T'_k} & \text{if} ~ k \in I \cap J \\
\text{otherwise} & \text{undefined}
\end{cases} \\
\end{array}
\]
\singlespacing

@@

Recall that routed selection and routed
branching behave in the same way as their ``non-routed''
counterparts -- the merging operator reflects this
similarity.

@@

Recall that well-formedness is a predicate defined 
\textit{solely} on the global type
in canonical MPST theory: a global type $G$ is well-formed
if a projection is defined for all its participants.

@@

We need to define the characteristics that $\role{s}$
must display in $G$ to prove that it is a
router. 
We formalise this as an \textit{inductively} defined relation,
$\centroid{G}{s}$, which reads \textit{``$\role{s}$ is a
centroid in $G$''}.
The intuition is that $\role{s}$ is at the centre
of all communication interactions.
We define what it means to be a \textit{centroid} in
\cref{def:centroid}.

@@

\begin{itemize}
\item 
\rulename{$\centroidop$-Comm}:
For direct communication, $\role{s}$ must
be a participant and the centroid of all continuations.

@@

We also assume that $G$ has no free type variables
and is guarded/contractive.
